%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FILE    : main.tex
% SUBJECT : Master document for the TISSEC 2014 paper.
% AUTHOR  : (C) Copyright 2014 by Peter C. Chapin and Christan Skalka
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prodmode,acmtissec]{acmsmall}

% --------
% Packages
% --------
\usepackage{fancyvrb}
\usepackage{stmaryrd}
\usepackage[english]{fp-autoref}
\usepackage{fp-frame}
\usepackage{soul}
\usepackage{mathpartir}
\usepackage{comment}
\usepackage{url}

% ------
% Macros
% ------
\include{macros}

% -----
% Figures made Chris' way 
% -----
\input{figures}

% To be filled in after the paper is accepted.
%\acmVolume{V}
%\acmNumber{N}
%\acmArticle{A}
%\acmYear{YYYY}
%\acmMonth{0}

\markboth{P. Chapin and C. Skalka}{RPC Authorization in Sensor Networks}

% Title portion
\title{SpartanRPC: Remote Procedure Call Authorization in Wireless Sensor Networks}
\author{PETER CHAPIN and CHRISTIAN SKALKA \affil{University of Vermont}}

\begin{abstract}
  We describe SpartanRPC, a secure middleware technology that supports
  cooperation between distinct security domains in wireless sensor
  networks. SpartanRPC extends nesC to provide a link-layer remote
  procedure call (RPC) mechanism, along with an enhancement of
  configuration wirings that allow specification of remote, dynamic
  endpoints. RPC invocation is secured via an authorization logic that
  enables servers to specify access policies, and requires clients to
  prove authorization. This mechanism is implemented using a
  combination of symmetric and public key cryptography. We report on
  benchmark testing of a prototype implementation, and on an
  application of the framework that supports secure collaborative use
  and administration of an existing WSN data gathering system.
\end{abstract}

\category{D.3.4}{Programming Languages}{Processors}[Compilers \and
Run-time environments]
\category{C.3}{Computer Systems Organization}{Special-Purpose and
  Application-Based Systems}[Real-time and embedded systems]
\category{C.2.4}{Computer-Communication Networks}{Distributed
  Systems}[Client/server]

\terms{Security, Languages}

\keywords{remote procedure call, sensor networks, trust management}

\acmformat{Peter Chapin and Christian Skalka. 2014. SpartanRPC: Remote
  Procedure Call Authorization in Wireless Sensor Networks}

\begin{document}

\begin{bottomstuff}
  Christian Skalka’s work was supported by a grant from the Air Force
  Office of Scientific Research Young Investigator Program (AFOSR YIP).

  Author's addresses: P. Chapin, (Current address) Computer Information
  Systems Department, Vermont Technical College; C. Skalka, Computer
  Science Department, University of Vermont.
\end{bottomstuff}

\maketitle

% ----------------
% Document content
% ----------------
\input{introduction}
\input{duties}
\input{dynamic-wires}
\input{security}
\input{extended-example}
\input{implementation}
\input{empirical-results}
\input{snowcloud}
\input{conclusion}

% ------------------------------
% Appendix and related materials
% ------------------------------

\bibliographystyle{ACM-Reference-Format-Journals}
\renewcommand{\showDOI}[1]{\unskip}
\bibliography{references-AccessControl,references-Meta,references-Security,references-WSN}

% History dates
\received{Month Year}{Month Year}{Month Year}

\end{document}
